---
title: "Time Series Analysis Project"
author: "Steven Dougherty, Bao Khanh Cu, Zherui Liang, Jiyang Huang"
date: "11/6/2019"
output: html_document
---
# Libraries in use for this project
```{r}
library(fpp2)
library(xts)
library(tseries)
library(fGarch)
library(rugarch)
library(Metrics)
library(forecast)
library(tsDyn)
this.dir = dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(this.dir)
```

## Setup:

Import Data set and transform into a time series

```{r}
SPclose<-read.csv("spx.csv",header = TRUE,sep = ",")
SPclose[,1]<-as.Date(SPclose[,1], "%d-%b-%y")
SP<- xts(SPclose[,-1], order.by=SPclose[,1])

## Not the cleanest method for getting the first plot point from each month, but is sufficient to account for all non-trading days early in the month.
SPmonthly<-c()
for(i in 1986:2017){
  for(j in 1:12){
    FirstDay<-FALSE
    if(FirstDay==FALSE){
   SPmonthlymin<-paste(as.character(i),"-",as.character(j),"-01",sep = "")
    minSPmonth<-which(index(SP)==SPmonthlymin)
    if(any(index(SP)==SPmonthlymin)){FirstDay<-TRUE}
    }
    if(FirstDay==FALSE){
   SPmonthlymin<-paste(as.character(i),"-",as.character(j),"-02",sep = "")
     minSPmonth<-which(index(SP)==SPmonthlymin)
     if(any(index(SP)==SPmonthlymin)){FirstDay<-TRUE}
    }
    if(FirstDay==FALSE){
   SPmonthlymin<-paste(as.character(i),"-",as.character(j),"-03",sep = "")
     minSPmonth<-which(index(SP)==SPmonthlymin)
     if(any(index(SP)==SPmonthlymin)){FirstDay<-TRUE}
    }
    if(FirstDay==FALSE){
   SPmonthlymin<-paste(as.character(i),"-",as.character(j),"-04",sep = "")
    minSPmonth<-which(index(SP)==SPmonthlymin)
    if(any(index(SP)==SPmonthlymin)){FirstDay<-TRUE}
    }
     if(FirstDay==FALSE){
   SPmonthlymin<-paste(as.character(i),"-",as.character(j),"-05",sep = "")
    minSPmonth<-which(index(SP)==SPmonthlymin)
    if(any(index(SP)==SPmonthlymin)){FirstDay<-TRUE}
    }
    SPmonthly<-c(SPmonthly,minSPmonth)
  }
}
i=2018
for(j in 1:6){
    FirstDay<-FALSE
    if(FirstDay==FALSE){
   SPmonthlymin<-paste(as.character(i),"-",as.character(j),"-01",sep = "")
    minSPmonth<-which(index(SP)==SPmonthlymin)
    if(any(index(SP)==SPmonthlymin)){FirstDay<-TRUE}
    }
    if(FirstDay==FALSE){
   SPmonthlymin<-paste(as.character(i),"-",as.character(j),"-02",sep = "")
     minSPmonth<-which(index(SP)==SPmonthlymin)
     if(any(index(SP)==SPmonthlymin)){FirstDay<-TRUE}
    }
    if(FirstDay==FALSE){
   SPmonthlymin<-paste(as.character(i),"-",as.character(j),"-03",sep = "")
     minSPmonth<-which(index(SP)==SPmonthlymin)
     if(any(index(SP)==SPmonthlymin)){FirstDay<-TRUE}
    }
    if(FirstDay==FALSE){
   SPmonthlymin<-paste(as.character(i),"-",as.character(j),"-04",sep = "")
    minSPmonth<-which(index(SP)==SPmonthlymin)
    if(any(index(SP)==SPmonthlymin)){FirstDay<-TRUE}
    }
     if(FirstDay==FALSE){
   SPmonthlymin<-paste(as.character(i),"-",as.character(j),"-05",sep = "")
    minSPmonth<-which(index(SP)==SPmonthlymin)
    if(any(index(SP)==SPmonthlymin)){FirstDay<-TRUE}
    }
    SPmonthly<-c(SPmonthly,minSPmonth)
  }
SPmonthly<-SP[SPmonthly,]

```

## Visualization of data:

Plotting the ACF, PACF, and plot of our timeseries

```{r}
autoplot(SPmonthly)+ylab("S&P 500 Index Price")+xlab("Date")
SP_log<-log(SPmonthly)
ggtsdisplay(SP_log)
```


## Transformation:
To make the data stationary, we will look at the log return.This is plotted along with its associated ACF and PACF below. The Augmented Dickey Fuller test returns a p-value of .01, indicating that this 


```{r}
SP_log<-as.ts(SP_log)
adf.test(SP_log)
# This data is not stationary, but a single difference will show stationarity
adf.test(diff(SP_log))
```
Split into test and training, where test data takes up the last 18 months.
```{r}
SP_train_log<-window(SP_log, end = 378)
SP_test_log<-window(SP_log, start = 379)
```


# Method 1: Time Series Linear Regression (ON log return)

```{r}
#data frame for tslm (log return)
#creating a time series that is indexed with the same dates as SPmonthly
SP_log_ret<-xts(rep(0,390),order.by =index(SPmonthly))
SP_log_ret[1]<-0
S<-SPmonthly%>%log()%>%diff()
SP_log_ret[2:390]<-na.omit(S)
## Importing our Predictors
X<-read.csv('predictors.csv') #predictors
X1<- xts(X[,-1], order.by=index(SPmonthly))
train_log_ret<-window(SP_log_ret,end=c("2017-06-01")) 
test_log_ret<-window(SP_log_ret, start=c("2017-07-03"))
SP_log_ret_train<-ts(data.frame(SP_INDEX=train_log_ret,X[1:378,]))
SP_log_ret_test<-ts(data.frame(SP_INDEX=test_log_ret,X[379:390,]))
```


```{r}
#model comparison
tslm(SP_INDEX ~ UNRATE+Usdollarindex+CPI+trend,data =SP_log_ret_train )%>%CV()
tslm(SP_INDEX ~ UNRATE+Usdollarindex+crudeoil+trend,data =SP_log_ret_train )%>%CV()# best
tslm(SP_INDEX ~ UNRATE+T_notes +crudeoil +trend,data =SP_log_ret_train )%>%CV()
tslm(SP_INDEX ~ CPI+T_notes +crudeoil +trend,data =SP_log_ret_train )%>%CV()
tslm(SP_INDEX ~ Usdollarindex+CPI+T_notes +trend,data =SP_log_ret_train )%>%CV()#best for logr
tslm(SP_INDEX ~ Usdollarindex+CPI +crudeoil +trend,data =SP_log_ret_train )%>%CV()
tslm(SP_INDEX ~ Usdollarindex+T_notes +crudeoil +trend,data =SP_log_ret_train )%>%CV()
tslm(SP_INDEX ~ UNRATE+T_notes +crudeoil +trend,data =SP_log_ret_train )%>%CV()
tslm(SP_INDEX ~ UNRATE+CPI +crudeoil +trend,data =SP_log_ret_train )%>%CV()
tslm(SP_INDEX ~ UNRATE+CPI+T_notes +trend,data =SP_log_ret_train )%>%CV()
tslm(formula = SP_INDEX ~ UNRATE + Usdollarindex + CPI + T_notes + 
    crudeoil + trend, data = SP_log_ret_train)%>%CV() # full model

```
## Checking Residuals for best model
```{r}
fit1<-tslm(SP_INDEX ~ Usdollarindex+CPI+T_notes +trend,data =SP_log_ret_train )
summary(fit1)
checkresiduals(fit1)
plot(fit1$fitted.values,SP_log_ret_train[,1],type='p',pch=16,cex=.2,main='Unemployment Rate + USD Index + T-note + CPI + trend') + abline(a=0,b=1,lwd=2,col='red')

```
```{r}
#forecast error of log return
newdata<-data.frame(DATE=SP_log_ret_test[,2],UNRATE=SP_log_ret_test[,3],Usdollarindex=SP_log_ret_test[,4],CPI=SP_log_ret_test[,5],T_notes=SP_log_ret_test[,6],crudeoil=SP_log_ret_test[,7])
fit1_forecast<-forecast(fit1, newdata = newdata)
forecast(fit1,newdata = data.frame(SP_log_ret_test),h=12)%>%autoplot()
```

```{r}
#Back transformation
tslmForecast<-rep(0,length(fit1_forecast$mean))
final_value<- SP_log[378]
tslmForecast[1] <- final_value+fit1_forecast$mean[1]
for (i in 2:length(fit1_forecast$mean)) { 
    tslmForecast[i] <- fit1_forecast$mean[i]+ tslmForecast[i-1]
}
#forecast error of transformed data
tslmForecast<-as.ts(ts(tslmForecast,start = 379, end = 390))
autoplot(as.ts(SP_log),include=50)+autolayer(tslmForecast)+ylim(7.5,8)+xlim(360,390)+ xlab("Month") +ylab("Log of Index Price")+theme(legend.position = "none")
```




# Method 2: ETS model:
# Application of ETS, using both ets function and manually with a for loop to find lowest AICc
```{r}

ETS_AICc_matrix<-matrix(10^8,nrow=2,ncol=5)
ETS_AICc_matrix[1,1]<-ets(SP_train_log,model = "ANN")$aicc
ETS_AICc_matrix[1,2]<-ets(SP_train_log,model = "AAN",damped = FALSE)$aicc
ETS_AICc_matrix[1,4]<-ets(SP_train_log,model = "AAN",damped = TRUE)$aicc
ETS_AICc_matrix[2,1]<-ets(SP_train_log,model = "MNN")$aicc
ETS_AICc_matrix[2,2]<-ets(SP_train_log,model = "MAN",damped = FALSE)$aicc
ETS_AICc_matrix[2,3]<-ets(SP_train_log,model = "MMN",damped = FALSE)$aicc
ETS_AICc_matrix[2,4]<-ets(SP_train_log,model = "MAN",damped = TRUE)$aicc
ETS_AICc_matrix[2,5]<-ets(SP_train_log,model = "MMN",damped = TRUE)$aicc
min(ETS_AICc_matrix)
ETS_AICc_matrix
ETS_AICc_matrix[1,2]
##AAN same choice manually as 
ETS_model<-ets(SP_train_log)

```


# Method 3: ARIMA model
## Application of ARIMA, doing a manual gridsearch for the optimal ARIMA(p,d,q) 
```{r}

d<-ndiffs(SP_train_log)
ARIMA_AICc_matrix<-matrix(0,nrow=6,ncol=6)
for(i in 1:6){
  for(j in 1:6){
    ARIMA_AICc_matrix[i,j]<-Arima(SP_train_log, order = c(i-1,d,j-1),include.drift = TRUE)$aicc
  }
}
ARIMA_AICc_matrix
min(ARIMA_AICc_matrix)
Arima_model<-auto.arima(SP_train_log, stepwise = FALSE, approximation = FALSE)
Arima_model$aicc
## ARIMA(1,1,1) is the best model (with draft)
Arima(SP_train_log, order = c(1,1,1),include.drift = TRUE)$aicc
```

##Check residulas of models with lowest AICc
ARIMA on full dataset, still have to split data into training and test 

```{r}
ETS_model<-ets(SP_train_log)
Arima_model<-auto.arima(SP_train_log, stepwise = FALSE, approximation = FALSE)
holtmod<-holt(SP_train_log,h=12,drift = TRUE)
ETS_model %>% forecast(h=40) %>% autoplot(PI=F)+autolayer(SP_test_log)+ xlab("Month") +ylab("Log of Index Price")+ylim(7.5,8)+xlim(360,390)+theme(legend.position = "none")
Arima_model %>% forecast(h=40) %>% autoplot(PI=F)+autolayer(SP_test_log)+ xlab("Month") +ylab("Log of Index Price")+ ylim(7.5,8)+xlim(360,390)+theme(legend.position = "none")

### Checking Residuals

checkresiduals(Arima_model)
ggAcf(Arima_model$residuals^2)
checkresiduals(ETS_model)

```


#Method 4: GARCH model
### GARCH Model
```{r}
##GARCH
fitgarch<-garchFit(~arma(1,1)+garch(1,1),diff(SP_train_log)[-1],trace=F)
summary(fitgarch)
garch.predict<- predict(fitgarch,12)
garch.predict$meanForecast
checkresiduals(fitgarch@residuals)
Box.test(fitgarch@residuals)



final_value<- SP_train_log[length(SP_train_log)]
garch.predict$meanForecast[1] <- final_value+garch.predict$meanForecast[1]
for (i in 2:12) { 
    garch.predict$meanForecast[i] <- garch.predict$meanForecast[i]+ garch.predict$meanForecast[i-1]
  
  }
```

```{r} 
plot.garch<-ts(garch.predict$meanForecast, start = 379, end = 390)
autoplot(as.ts(plot.garch))+ xlab("Month")+autolayer(as.ts(SP_log)) + ylab("Log of Index Price")+ylim(7.5,8)+xlim(360,390)+theme(legend.position = "none")
```

#Method 5: Neural network


```{r}
set.seed(1)
neural<- nnetar(SP_train_log, p = 12, size = 6)
neural.forecast <- forecast(neural, h = 12)
autoplot(neural.forecast) +autolayer(as.ts(SP_log)) + xlab("Month") +ylab("Log of Index Price")+ylim(7.5,8)+xlim(360,390)+theme(legend.position = "none")

```

### Neural Network alternative
```{r}
set.seed(1)
nnet<- nnetTs(SP_train_log, m=8, size=6)
nnet.predict<-predict(nnet,n.ahead=12)
autoplot(nnet.predict) +autolayer(as.ts(SP_log)) + xlab("Month") +ylab("Log of Index Price")+ylim(7.5,8.25)+xlim(360,390)+theme(legend.position = "none")
AIC(nnet)

```

#Method 6: SETAR model
```{r}

fit.setar<-setar(train_log_ret,m=2,thDelay = 1)
fit.setar
pred.setar <- predict(fit.setar,n.ahead=12)
pred.setar[1]<-pred.setar[1]+SP_train_log[378]
for (i in 2:12){
  pred.setar[i]<-pred.setar[i]+pred.setar[i-1]
}
pred.setar
sum((pred.setar -  SP_test_log)^2) 
AIC(fit.setar)
checkresiduals(fit.setar$residuals)
Box.test(fit.setar$residuals,lag=10,type='Ljung')
autoplot(SP_log)+autolayer(pred.setar)+ xlab("Month") +ylab("Log of Index Price")+ylim(7.5,8)+xlim(360,390)+theme(legend.position = "none")

```

## Checking AIC
```{r}

#ETS
AIC_ETS<-ETS_model$aicc
AIC_ETS

#ARIMA
AIC_Arima<-Arima_model$aicc
AIC_Arima

#GARCH
AIC_Garch<-fitgarch@fit$ics[1]*length(SP_train_log)
AIC_Garch

# #Neural
AIC_NN<-AIC(nnet)
AIC_NN

## SETAR
AIC_SETAR<-AIC(fit.setar)
AIC_SETAR

```
### Checking Fit For All Models
```{r}
# TSLM
# forecast::accuracy(tslmForecast,SP_test_log) %>% round(4)
MSE_TSLM<-sum((tslmForecast-SP_test_log)^2)
MSE_TSLM

#ETS
# ETS_model %>% forecast(h=12) %>%forecast::accuracy(SP_test_log) %>% round(4)
fETS<-forecast(ETS_model,h=12)
MSE_ETS<-sum((fETS$mean-SP_test_log)^2)
MSE_ETS

#ARIMA
# Arima_model %>% forecast(h=12) %>% forecast::accuracy(SP_test_log) %>% round(4)
fArima<-forecast(Arima_model,h=12)
MSE_Arima<-sum((fArima$mean-SP_test_log)^2)
MSE_Arima

#GARCH
# forecast::accuracy(ts(SP_test_log, start=379, end=390),plot.garch)%>%round(4)
MSE_Garch<-sum((plot.garch-SP_test_log)^2)
MSE_Garch

# #Neural
# forecast::accuracy(neural.forecast, SP_test_log)%>% round(4)
MSE_NN<-sum((nnet.predict-SP_test_log)^2) 
MSE_NN

## SETAR
MSE_SETAR<-sum((pred.setar-SP_test_log)^2)
MSE_SETAR

```
## SETAR has the lowest test errors of the three models presented here.


### Graph of Test Data with all models
```{r}
dat.predict<-data.frame(Time<-index(SPmonthly[379:390]),actual<-SP_test_log,TSLM<-tslmForecast,ETS<-fETS$mean,ARIMA<-fArima$mean,NN<-nnet.predict,GARCH<-plot.garch,SETAR<-pred.setar)

ggplot(dat.predict, mapping=aes(x=Time))+geom_line(mapping=aes(y=actual))+geom_line(mapping=aes(y=TSLM))+geom_line(mapping=aes(y=ETS))+geom_line(mapping=aes(y=ARIMA))+geom_line(mapping=aes(y=NN))+geom_line(mapping=aes(y=GARCH))+geom_line(mapping=aes(y=SETAR))
```




